---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}
h2 {
  font-weight: bold;
  text-decoration: underline;
}
h3 {
  font-weight: bold;
}
```

## How Does Heart Rate Vary Throughout the Day?

This dynamic report aims to investigate how heart rate changes throughout the day including which activities were being done and what time of day it was.

```{r}
library(magick)

image_read(path = "https://c.tenor.com/Xx2YeqghQtkAAAAd/tenor.gif") %>%
  image_scale(200)
```

```{r}
library(tidyverse)

logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTCIlpUvabeVRVhMybVDz--d9tMtNJUHnrwwEnGMnvOZuP_SdXyPq46ip4oYhvKMuiRnSiRHw3oYnBd/pub?output=csv")
latest_data <- logged_data %>%
  rename(date_and_time = 1,
         heart_rate_bpm = 2,
         time_of_day = 3,
         activity = 4)


#----------------------Exploration----------------------

#Creating new dataframe for average heart rates across different times of the day

time_of_day = c("Morning (7am - 11am)",                                                               #Variable for different times of the day
                "Afternoon (12pm - 4pm)",
                "Evening (5pm - 9pm)",
                "Night (9pm - 1am)")

hr_means <- c(mean(latest_data$heart_rate_bpm[latest_data$time_of_day=="Morning (7am - 11am)"]),      #Variable for mean heart rates across different times of the day
              mean(latest_data$heart_rate_bpm[latest_data$time_of_day=="Afternoon (12pm - 4pm)"]),
              mean(latest_data$heart_rate_bpm[latest_data$time_of_day=="Evening (5pm - 9pm)"]),
              mean(latest_data$heart_rate_bpm[latest_data$time_of_day=="Night (9pm - 1am)"]))

time_means <- data.frame(time_of_day, hr_means)                                                       #Putting the new variables into a new dataframe


#Creating new dataframe for average heart rates across grouped activities

eat_drink <- mean(c(latest_data$heart_rate_bpm[latest_data$activity=="Lunch"],            #Variable for activities I consider eating/drinking
               latest_data$heart_rate_bpm[latest_data$activity=="Drinking coffee"],
               latest_data$heart_rate_bpm[latest_data$activity=="Dinner"],
               latest_data$heart_rate_bpm[latest_data$activity=="Breakfast"]))

study <- mean(c(latest_data$heart_rate_bpm[latest_data$activity=="Walking home"],         #Variable for activities I consider study-related
           latest_data$heart_rate_bpm[latest_data$activity=="Studying"],
           latest_data$heart_rate_bpm[latest_data$activity=="Study"],
           latest_data$heart_rate_bpm[latest_data$activity=="Going to class"]))

chores <- mean(c(latest_data$heart_rate_bpm[latest_data$activity=="Laundry"],             #Variable for activities I consider chores
            latest_data$heart_rate_bpm[latest_data$activity=="Groceries"],
            latest_data$heart_rate_bpm[latest_data$activity=="Dishes"],
            latest_data$heart_rate_bpm[latest_data$activity=="Cooking"],
            latest_data$heart_rate_bpm[latest_data$activity=="Buying lunch"],
            latest_data$heart_rate_bpm[latest_data$activity=="Buying dinner"],
            latest_data$heart_rate_bpm[latest_data$activity=="Brushing teeth"]))

recreation <- mean(c(latest_data$heart_rate_bpm[latest_data$activity=="Youtube"]))        #Variable for activities I consider recreation

activity_means <- c(eat_drink, study, chores, recreation)                                 #Putting the new variables into a new dataframe
activity_type <- c("Eating/Drinking", "Study", "Chores", "Recreation")
grouped_activity <- data.frame(activity_means, activity_type)

```

### Heart Rate by Time

This first graph shows us the average heart rate at each time of day separated into categories (Morning, Afternoon, Evening, and Night).

```{r}

#Mean heart rates by time of day
time_means %>%
  ggplot(aes(x = time_of_day, y = hr_means)) +
    geom_bar(stat = "identity", fill = "#84CDE5") +
  labs(title = "Mean Heart Rates by Time of Day",
       subtitle = "Comparing which time of the day has the highest average heart rate",
       caption = "Subject: William Tess",
       x = "Time of Day",
       y = "Mean Heart Rate (bpm)")

```

We can see from this graph that average heart rate was highest in the afternoon (`r hr_means[2]` bpm), followed by the evening (`r hr_means[3]` bpm), then night (`r hr_means[4]` bpm), then the morning where average heart rate was lowest (`r hr_means[1]` bpm).

### Heart Rate by Activity

This second graph describes average heart rate depending on the type of activity being done. Activities are categorised into being related to chores, eating/drinking, studying, and recreation.

```{r}
#Mean heart rates by activity type
grouped_activity %>%
  ggplot(aes(y = activity_means, x = activity_type)) +  
    geom_bar(stat = "identity", fill = "#84CDE5") +
  labs(title = "Mean Heart Rates by Activity Type",
      subtitle = "Comparing which type of activity has the highest average heart rate",
      caption = "Subject: William Tess",
      x = "Activity Type",
      y = "Mean Heart Rate (bpm)")

```

The activity type which caused the highest average heart rate was chores (`r activity_means[3]` bpm), followed by eating/drinking (`r activity_means[1]` bpm), then studying (`r activity_means[2]` bpm), and lastly recreation which caused the lowest average heart rate (`r activity_means[4]` bpm).
